---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

An introduction of the data and a description of the trends/books/items you are choosing to analyze (and why!)

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

These will likely be calculated using your DPLYR skills, answering questions such as:

- What is the average number of checkouts for each item?
- What is the month or year with the most/least checkouts for a book that you're interested in?
- What is the month or year with the most/least checkouts for ebooks?
- How has the number of print book checkouts changed over time?

Feel free to calculate and report values that you find relevant.

### The Dataset

- Who collected/published the data?
  - The data is published by the Seattle Public Library.
- What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?
  - The parameters of the data are UsageClass, CheckoutType, MaterialType, CheckoutYear, CheckoutMonth, Checkouts, Title, ISBN, Creator, Subjects, Publisher, and PublicationYear.
- How was the data collected or generated?  
  - The data is generated through SPL checkout data which comes from multiple historical and current sources. Particular for historical physical item checkouts, the data is based on Legrady artwork data archives for checkouts between April 2005 and September 30, 2016. From forward October 2016 to present day, it's from Horizon ILS.
- Why was the data collected?  
  - The data was collected to record trends of library information. Trends such as noting popular interest of what the communities they serve are checking out. 
- What, if any, ethical questions do you need to consider when working with this data?  
  - How does the library differentiate between duplicate entries when some books don't contain an ISBN?
- What are possible limitations or problems with this data?   (at least 200 words)
  - One limitations with the dataset is that there is inconsistency in the Creator column, where the name of the creator's full name can be printed backwards - such as last name coming before the first. In some cases, alongside the creator's name, a printed year is associated with their name.
  - Another limitation with the data set is that some of the values has missing data. For instance, ISBN and creator has a few empty spaces within the dataset.
  - Another limitation would be is that it doesn't include any information of the belonging SPL branch the item is checked out from. This kind of data would be useful to determine patterns among library locations across the Seattle area for specific items that are checked out.

### First Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, warning = FALSE, code = readLines("chart_one.R")}
```

### Second Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show another trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, code = readLines("chart2_example.R")}
```

### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more. Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title
- You need a legend for your different line colors and a clear legend title

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, code = readLines("Chart3.R")}
```
